---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# featherframe

Sometimes with `reticulate` you import Python libraries that require using 
Pandas DataFrame and Series objects. Currently there is support for Numpy 
Arrays in `reticulate`, but no support for these other objects. 

`feather` is a great package that allows you to write R `tibble`s to disk, and read them
into Python as `DataFrames`, and vice versa, with type support. Combining `reticulate` and `feather` results in a way to more easily convert R objects to Pandas DataFrame
and Series objects for use with other libraries you might import with `reticulate`.
It is actually surprisingly fast too!

## Installation

You can install `featherframe` from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("DavisVaughan/featherframe")
```

Note that you will have to ensure that any Python library you want to use
(like in the examples below) is installed on your machine. You also will need
to ensure that `reticulate` is finding the correct version of Python, and how
to do all of that is nicely outlined for you in their [vignette](https://rstudio.github.io/reticulate/articles/versions.html). Personally
I have set the `RETICULATE_PYTHON` environment variable.

## Example - Conversion

`tibble` to `Pandas DataFrame` and back.

```{r example, message=FALSE}
library(dplyr)
library(featherframe)
data("iris")

# data.frame to tibble
iris <- iris %>%
  as_tibble()

# tibble to Pandas DataFrame
iris_pd <- iris %>%
  as_pandas_DataFrame()

iris_pd

# Pandas DataFrame to tibble
iris_tib <- iris_pd %>%
  as_tibble()

iris_tib

# It works!
identical(iris, iris_tib)
```

## Example - The Python pyfolio library

The original motivation for creating this mini-package is to use to [pyfolio](https://github.com/quantopian/pyfolio) library from Quantopian right from
R. It provides some neat financial analysis functions I was interested in testing,
but requires you to pass in Pandas Series and DataFrame objects.

Specifically, I am going to demonstrate some functionality that requires a Pandas Series
of daily returns.

```{r, message=FALSE}
library(reticulate)
library(featherframe)
library(tibbletime) # just for a test data set
library(lubridate)
library(dplyr)

# Use reticulate to import pyfolio
pyfolio <- import("pyfolio")

# Get the FB stock dataset from tibbletime
data(FB)
FB

# There is an issue with Dates becoming POSIXct so let's avoid that
FB <- mutate(FB, date = as_datetime(date))

# Calculate returns
FB_returns <- FB %>%
  mutate(returns = adjusted / lag(adjusted) - 1) %>%
  select(date, returns) %>%
  na.omit()

# Convert to Pandas Series, specifying the index column
FB_returns_pd <- FB_returns %>%
  as_pandas_Series(index = date)

FB_returns_pd

# Use the rolling_volatility function from the current pip version of pyfolio
rolling_vol_pd <- pyfolio$timeseries$rolling_volatility(FB_returns_pd, rolling_vol_window = 10L)
rolling_vol_pd

# Go back to R now
rolling_vol_pd %>%
  as_tibble() %>%
  rename(returns_rolling_vol = returns)
```

The idea is that we can wrap these functions up to take care of all the conversion
for us. I believe this is essentially what the `keras` package does with the Python `keras`
library, but in a more efficient way than using `feather`.

```{r}
pyfolio_rolling_volatility <- function(x, rolling_vol_window, ...) {
  pyfolio <- import("pyfolio")
  stopifnot(is.integer(rolling_vol_window))
  
  as_pandas_Series(x, ...) %>%
    pyfolio$timeseries$rolling_volatility(rolling_vol_window) %>%
    as_tibble()
}

pyfolio_rolling_volatility(FB_returns, 10L, index = date)
```


## Known issues

Currently `Date` columns are translated to `datetime64[ns]` and then back as
`POSIXct` with a local-specific timezone. This is...undesirable and should be
fixed on the `feather` side of things.

```{r}
ex <- tibble(date_col = as.Date("2018-01-01"))
ex_pd <- as_pandas_DataFrame(ex)
ex_pd

as_tibble(ex_pd)
```

Ordered factors do not stay ordered (but remain factors) when going back and forth.

```{r}
ex_ordered_fac <- tibble(fct = factor(c("low", "high", "med"), 
                                      levels = c("low", "med", "high"), 
                                      ordered = TRUE))
ex_ordered_fac

as_tibble(as_pandas_DataFrame(ex_ordered_fac))
```

Integer columns with `NA` values are coerced to numeric (double) column because
Pandas does not have support for `NaN` values in integer columns.

```{r}
ex_int <- tibble(int_col = c(1L, 2L, 3L))
ex_int

ex_int_NA <- tibble(int_col = c(1L, 2L, 3L, NA))
ex_int_NA

as_tibble(as_pandas_DataFrame(ex_int))
as_tibble(as_pandas_DataFrame(ex_int_NA))
```

Rownames are not preserved in the translation from R to Python.
